---
title: "workflow2"
author: "Artur Jacques Nürnberg"
date: "8/4/2020"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)

sales <- read_csv("sales2019.csv")
head(sales)
```

How big is the dataset? What are the column names, and what do they seem to represent?

```{r}
dim(sales)
```
The dataset has 5000 lines and 5 columns, `date` that represents the date that 
was purchase, `user_submitted_review` that is the review, `title` that is the 
title of the book, `total_purchased` that represents the total purchased and
`customer_type` that represents the customer type.


What are the types of each of the columns?

```{r}
for (name in colnames(sales)){
  print(paste(name, "is of the type", typeof(sales[[name]])))
}
```
That being said, do any of the columns have missing data? If so, make a note of 
this.

```{r}
for (name in colnames(sales)){
  print(paste(name, 'is missing' ,sum(is.na(sales[[name]]))))
}
```
Remove all rows in the dataset that have an NA value for the 
user_submitted_review column.

```{r}
sales_review <- filter(sales, !is.na(sales$user_submitted_review))
head(sales_review)
```

Using the remaining rows that have data, calculate the average number of books 
purchased on an order.

```{r}
avg_purchase <- round(mean(sales_review$total_purchased, na.rm = T))
avg_purchase
```
Fill all of the missing values in total_purchased with the average value you 
calculated.

```{r}
sales_review_filled <- sales_review %>% 
  mutate(total_purchased = if_else(is.na(total_purchased), avg_purchase,total_purchased))
head(sales_review_filled)
```
Create a new column that indicates if the review is positive or not

```{r}
unique(sales_review_filled$user_submitted_review)

sales_review_filled <- sales_review_filled %>% mutate(
  is_positive = case_when(
    str_detect(user_submitted_review, 'OK') ~ T,
    str_detect(user_submitted_review, 'okay') ~ T,
    str_detect(user_submitted_review, 'Hated') ~ F,
    str_detect(user_submitted_review, 'learned') ~ T,
    str_detect(user_submitted_review, 'Awesome') ~ T,
    str_detect(user_submitted_review, 'Never') ~ T,
    str_detect(user_submitted_review, 'not recommend') ~ F,
    TRUE ~ FALSE
  )
)

sales_review_filled %>% select("user_submitted_review", "is_positive") %>% unique()
```

Perform the proper conversion of the date column, so that it actually represents
a date and time
Create a new grouping column using the mutate() function that would help 
distinguish between sales that happen before July 1, 2019 and sales that happen 
after this date.

```{r}
sales_review_filled_dated <- sales_review_filled %>% 
  mutate(
    date = mdy(date),
    before_first_july = date < ymd('2019-7-1')
  )
sales_review_filled_dated %>% select('date', 'before_first_july')
```
Create a summary table that compares the number of books purchased before 
July 1, 2019 to after.

```{r}
sales_review_filled_dated %>% 
  group_by(before_first_july, customer_type) %>%
  summarise( total_reviewed= n(), positive_review = sum(is_positive)/n(), 
             total_purchased = sum(total_purchased))
```

Bussines purhcase more than individuals in the two periods.




