---
title: "Analyzing Forest Fire Data"
author: "Artur Jacques Nürnberg"
date: "8/6/2020"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
```

```{r}
fire <- read_csv('forestfires.csv') 
head(fire)
```

X: X-axis spatial coordinate within the Montesinho park map: 1 to 9
Y: Y-axis spatial coordinate within the Montesinho park map: 2 to 9
month: Month of the year: 'jan' to 'dec'
day: Day of the week: 'mon' to 'sun'
FFMC: Fine Fuel Moisture Code index from the FWI system: 18.7 to 96.20
DMC: Duff Moisture Code index from the FWI system: 1.1 to 291.3
DC: Drought Code index from the FWI system: 7.9 to 860.6
ISI: Initial Spread Index from the FWI system: 0.0 to 56.10
temp: Temperature in Celsius degrees: 2.2 to 33.30
RH: Relative humidity in percentage: 15.0 to 100
wind: Wind speed in km/h: 0.40 to 9.40
rain: Outside rain in mm/m2 : 0.0 to 6.4
area: The burned area of the forest (in ha): 0.00 to 1090.84

During which months are forest fires most common?

```{r}
burned <- fire %>% group_by(month) %>% summarise(total_fires = n()) %>%
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may",
                                          "jun", "jul", "aug", "sep", "oct",
                                          "nov", "dec")
                        )
         )
ggplot(data = burned) +
  aes(x = month, y = total_fires) +
  geom_bar(stat = 'identity') +
  labs(title = 'Fires on the months', x = 'Month', y = 'Fires')
```
The months that have more fires are august and september that are the end of the
summer.


On which days of the week are forest fires most common?

```{r}
burned <- fire %>% group_by(day) %>% summarise(total_fires = n()) %>% 
  mutate(day = factor(day, levels = c('sun', 'mon', 'tue', 'wed', 'thu', 'fri','sat')
                        )
         )
ggplot(data = burned) +
  aes(x = day, y = total_fires) +
  geom_bar(stat = 'identity') +
  labs(title = 'Fires on the months', x = 'Month', y = 'Fires')
```
Close to the weekend that is more fires that another days. Maybe because of a
human factor.

```{r}
fire <- fire %>% mutate(
  month = factor(
    month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug",
                      "sep", "oct","nov", "dec"
                      )
    ),
  day = factor(day, levels = c('sun', 'mon', 'tue', 'wed', 'thu', 'fri','sat'))
  )
```


Create box plots to visualize the distribution of the following variables by month and by day of the week:
FFMC
DMC
DC
ISI
temp
RH
wind
rain

```{r}
y_vector = c('FFMC', 'DMC', 'DC', 'ISI', 'temp', 'RH', 'wind', 'rain')

month_graph <- function(y_var){
  ggplot(data = fire) +
  aes_string(x = 'month', y = y_var) +
  geom_boxplot()
}

map(y_vector, month_graph)
```
```{r}
day_graph <- function(y_var){
  ggplot(data = fire) +
  aes_string(x = 'day', y = y_var) +
  geom_boxplot()
}

map(y_vector, day_graph)
```
Let's create scatter plots to see what can be learned about relationships between the area burned by a forest fire and the following variables:

FFMC: Fine Fuel Moisture Code index from the FWI system: 18.7 to 96.20
DMC: Duff Moisture Code index from the FWI system: 1.1 to 291.3
DC: Drought Code index from the FWI system: 7.9 to 860.6
ISI: Initial Spread Index from the FWI system: 0.0 to 56.10
temp: Temperature in Celsius degrees: 2.2 to 33.30
RH: Relative humidity in percentage: 15.0 to 100
wind: Wind speed in km/h: 0.40 to 9.40
rain: Outside rain in mm/m2 : 0.0 to 6.4


```{r}
day_scatter_graph <- function(x_var){
  ggplot(data = fire) +
  aes_string(x =x_var, y = 'area') +
  geom_point()
}

map(y_vector, day_scatter_graph)
```
when rains rarely fires burn large areas, increase of FFMC cause a larger area
burned.
