---
title: "Designing and Creating a Database"
author: "Artur Jacques Nürnberg"
date: "8/21/2020"
output: html_document
---

```{r}
library(tidyverse)
library(RSQLite)
```


```{r}
game_log <- read_csv('game_log.csv')

head(game_log)
```
```{r}
park_codes <- read_csv('park_codes.csv')

head(park_codes)
```
```{r}
person_codes <- read_csv('person_codes.csv')

head(person_codes)
```
```{r}
team_codes <- read_csv('team_codes.csv')

head(team_codes)
```
```{r}
appearance_type <- read_csv('appearance_type.csv')

head(appearance_type)
```


Use dbWriteTable() to create tables for each of our csv files into a new SQLite 
database, mlb.db:
The table name should be the same as each of the CSV filename without the 
extension, eg game_log.csv should be imported to a table called game_log.

```{r}
conn <- dbConnect(SQLite(), 'mlb.db')

dbWriteTable(conn = conn,
             name = 'game_log',
             value = game_log,
             header = TRUE)

dbWriteTable(conn = conn,
             name = 'park_codes',
             value = park_codes,
             header = TRUE)

dbWriteTable(conn = conn,
             name = 'person_codes',
             value = person_codes,
             header = TRUE)

dbWriteTable(conn = conn,
             name = 'team_codes',
             value = team_codes,
             header = TRUE)

dbWriteTable(conn = conn,
             name = 'appearance_type',
             value = appearance_type,
             header = TRUE)

dbDisconnect(conn)
```

```{r}
run_action <- function(query){
  conn <- dbConnect(SQLite(), 'mlb.db')
  result <- dbExecute(conn, query)
  dbDisconnect(conn)
  return(result)
}

run_query <- function(query){
  conn <- dbConnect(SQLite(), 'mlb.db')
  result <- dbGetQuery(conn, query)
  dbDisconnect(conn)
  return(result)
}

show_tables <- function(){
  run_query("SELECT name FROM sqlite_master WHERE type='table'")
}
```

Create a new column in the game_log table called game_id:
Use SQL string concatenation to update the new columns with a unique ID using 
the Retrosheet format outlined above.
Once you write the query, you can use dbExecute() to perform the query

```{r}
create_column <-"
ALTER TABLE game_log
ADD COLUMN game_id TEXT;
"
run_action(create_column)

alter_column <- "
UPDATE game_log
SET game_id = CAST(date AS INTEGER) || h_name || CAST(number_of_game AS INT);"

run_action(alter_column)

run_query('SELECT game_id FROM game_log')
```
Looking at the various files, look for opportunities to normalize the data and 
record your observations in a markdown cell.

removing column repetition

```{r}
remove_columns <- "

"
```


Create the person table with columns and primary key as shown in the schema 
diagram.
Select the appropriate type based on the data.
Insert the data from the person_codes table.
Write a query to display the first few rows of the table.

```{r}
create_table_person <- "
CREATE TABLE person(
  person_id TEXT PRIMARY KEY,
  first_name TEXT,
  last_name TEXT
)
"

run_action(create_table_person)

insert_person_data <-"
INSERT INTO person
SELECT 
  id,
  first,
  last
FROM PERSON_CODES
"

run_action(insert_person_data)

run_query('SELECT * FROM person LIMIT 5')
```

Create the park table with columns and primary key as shown in the schema 
diagram.
Select the appropriate type based on the data
Insert the data from the park_codes table.
Write a query to display the first few rows of the table.

```{r}
create_table_park <-"
CREATE TABLE park(
  park_id TEXT PRIMARY KEY,
  name TEXT,
  city TEXT,
  state VARCHAR(2),
  notes TEXT
)
"

run_action(create_table_park)

insert_park_data <-"
INSERT INTO park
  SELECT
    park_id,
    name,
    city,
    state,
    notes
  FROM park_codes
"
run_action(insert_park_data)
run_query('SELECT * FROM park LIMIT 5')
```

Create the league table with columns and primary key as shown in the schema
diagram.
Select the appropriate type based on the data.
Insert the data manually based on your research on the names of the six league 
IDs.
Write a query to display the table.

```{r}
create_table_league <-"
CREATE TABLE league(
  league_id INTEGER PRIMARY KEY,
  name varchar(2)
);
"

run_action(create_table_league)

insert_league_data <-"
INSERT INTO league
SELECT
  ROW_NUMBER() OVER(ORDER BY COUNT(league) DESC) AS league_id,
  league
FROM team_codes
GROUP BY league
"

run_action(insert_league_data)

run_query('SELECT * FROM league LIMIT 5')
```
Create the team table with columns, primary key, and foreign key as shown in the 
schema diagram.
Select the appropriate type based on the data.
Insert the data from the team_codes table.
Write a query to display the first few rows of the table.

```{r}
create_table_team <-"
CREATE TABLE team(
  team_id TEXT PRIMARY KEY,
  league_id INTEGER,
  city TEXT,
  nickname TEXT,
  franch_id TEXT,
  FOREIGN KEY (league_id) REFERENCES league(league_id)
);
"

run_action(create_table_team)

insert_team_data <-"
INSERT INTO team
SELECT
  t.team_id,
  l.league_id,
  t.city,
  t.nickname,
  franch_id
FROM team_codes t
LEFT JOIN league l ON l.name = t.league
GROUP BY t.team_id;
"

run_action(insert_team_data)

run_query("SELECT * FROM team")
```

Create the game table with columns, primary key, and foreign key as shown in the
schema diagram.
Select the appropriate type based on the data.
Insert the data from the game_log table.
Write a query to display the first few rows of the table.

```{r}
create_table_game <-"
CREATE TABLE game(
  game_id TEXT PRIMARY KEY,
  date INTEGER,
  number_of_game INTEGER,
  park_id TEXT,
  length_outs INTEGER,
  day TEXT,
  completion INTEGER,
  forfeit TEXT,
  protest INTEGER,
  attendance INTEGER,
  lenght_minutes INTEGER,
  additional_info TEXT,
  acquisition_info TEXT,
  FOREIGN KEY (park_id) REFERENCES park(park_id)
);
"

run_action(create_table_game)

insert_game_data <-"
INSERT INTO game
SELECT
  game_id,
  date,
  number_of_game,
  park_id,
  length_outs,
  day_night,
  completion,
  forefeit,
  protest,
  attendance,
  length_minutes,
  additional_info,
  acquisition_info
FROM game_log
"

run_action(insert_game_data)

run_query('SELECT * FROM game LIMIT 5')
```
Create the team_appearance table with columns, primary key, and foreign keys as 
shown in the schema diagram.
Select the appropriate type based on the data.
Insert the data from the game_log table, using a UNION clause to combine the 
data from the column sets for the home and away teams.
Write a query to verify that your data was inserted correctly.

```{r}
create_table_appearance <-"
CREATE TABLE team_appearance(
  team_id TEXT,
  game_id TEXT,
  home INTEGER,
  league_id INTEGER,
  score NUMERIC,
  line_score TEXT,
  at_bats NUMERIC,
  hits NUMERIC,
  doubles NUMERIC,
  triples NUMERIC,
  homeruns NUMERIC,
  rbi NUMERIC,
  sacrifice_hits NUMERIC,
  sacrifice_flies NUMERIC,
  hit_by_pitch NUMERIC,
  walks NUMERIC,
  intentional_walks NUMERIC,
  strikeouts NUMERIC,
  stolen_bases NUMERIC,
  caught_stealing NUMERIC,
  grounded_into_double NUMERIC,
  first_catcher_interference INTEGER,
  left_on_base NUMERIC,
  pitchers_used NUMERIC,
  individual_earned_runs NUMERIC,
  team_earned_runs NUMERIC,
  wild_pitches NUMERIC,
  balks NUMERIC,
  putouts NUMERIC,
  assists NUMERIC,
  errors NUMERIC,
  passed_balls NUMERIC,
  double_plays NUMERIC,
  triple_plays NUMERIC,
  FOREIGN KEY (league_id) REFERENCES league(league_id),
  PRIMARY KEY (team_id, game_id)
);
"

run_action(create_table_appearance)

insert_appearance_data <- ("
INSERT INTO team_appearance
  SELECT
    h_name,
    game_id,
    1 AS home,
    h_league,
    h_score,
    h_line_score,
    h_at_bats,
    h_hits,
    h_doubles,
    h_triples,
    h_homeruns,
    h_rbi NUMERIC,
    h_sacrifice_hits,
    h_sacrifice_flies,
    h_hit_by_pitch,
    h_walks,
    h_intentional_walks,
    h_strikeouts,
    h_stolen_bases,
    h_caught_stealing,
    h_grounded_into_double,
    h_first_catcher_interference,
    h_left_on_base,
    h_pitchers_used,
    h_individual_earned_runs,
    h_team_earned_runs,
    h_wild_pitches,
    h_balks,
    h_putouts,
    h_assists,
    h_errors,
    h_passed_balls,
    h_double_plays,
    h_triple_plays
  FROM game_log
  
  UNION
  
  SELECT
    v_name,
    game_id,
    0 AS home,
    v_league,
    v_score,
    v_line_score,
    v_at_bats,
    v_hits,
    v_doubles,
    v_triples,
    v_homeruns,
    v_rbi NUMERIC,
    v_sacrifice_hits,
    v_sacrifice_flies,
    v_hit_by_pitch,
    v_walks,
    v_intentional_walks,
    v_strikeouts,
    v_stolen_bases,
    v_caught_stealing,
    v_grounded_into_double,
    v_first_catcher_interference,
    v_left_on_base,
    v_pitchers_used,
    v_individual_earned_runs,
    v_team_earned_runs,
    v_wild_pitches,
    v_balks,
    v_putouts,
    v_assists,
    v_errors,
    v_passed_balls,
    v_double_plays,
    v_triple_plays
  FROM game_log
")

run_action(insert_appearance_data)


run_query('SELECT * FROM team_appearance LIMIT 5')
```
Create the person_appearance table with columns, primary key, and foreign keys 
as shown in the schema diagram.
Select the appropriate type based on the data.
Insert the data from the game_log table, using UNION clauses to combine the data
from the columns for managers, umpires, pitchers, and awards.
Use a loop with string formatting to insert the data for offensive and defensive
positions from the game_log table.
Write a query to verify that your data was inserted correctly.

```{r}
create_table_person_appearance <- "
CREATE TABLE person_appearance(
  apperance_id INTEGER PRIMARY KEY AUTOINCREMENT,
  person_id TEXT,
  team_id TEXT,
  game_id TEXT,
  appearance_type_id TEXT,
  FOREIGN KEY (person_id) REFERENCES person(person_id),
  FOREIGN KEY (team_id) REFERENCES team_appearance(team_id),
  FOREIGN KEY (game_id) REFERENCES game(game_id),
  FOREIGN KEY (appearance_type_id) REFERENCES appearance_type(appearance_type_id)
)
"

run_action(create_table_person_appearance)

for (letter in c('h', 'v')){
  for (num in 1:9){
    insert_person_appearance_data <-"
    INSERT INTO person_appearance (
      game_id,
      team_id,
      person_id,
      appearance_type_id
    )
      SELECT
        game_id,
        %s_name,
        %s_player_%f_id,
        'O%f'
      FROM game_log
      WHERE %s_player_%f_id IS NOT NULL
    
    UNION
    
      SELECT
        game_id,
        %s_name,
        %s_player_%f_id,
        'D' || cast(%s_player_%f_def_pos AS iNT)
      FROM game_log
      WHERE %s_player_%f_id IS NOT NULL;
    "
    insert_person_appearance_data <- gsub('%s',
                                          letter,
                                          insert_person_appearance_data,
                                          fixed = TRUE)
    insert_person_appearance_data <- gsub('%f',
                                          num,
                                          insert_person_appearance_data,
                                          fixed = TRUE)
    
    run_action(insert_person_appearance_data)
  }
}


run_query("
SELECT * FROM person_appearance LIMIT 5
")
```
Drop the tables we created to hold our unnormalized data:
game_log.
park_codes.
team_codes.
person_codes.

```{r}
run_action('
           DROP TABLE game_log;
           DROP TABLE park_codes;
           DROP TABLE team_codes;
           DROP TABLE person_codes;
           ')
```

