---
title: "Business Questions using SQL"
author: "Artur Jacques Nürnberg"
date: "8/19/2020"
output: html_document
---

Import the RSQLite and DBI libraries.

```{r}
library(RSQLite)
library(DBI)
library(ggplot2)
library(dplyr)
```

Create a run_query() function that takes in a SQL query as an argument and 
returns the result in a dataframe.

```{r}
run_query <- function(query){
  conn <- dbConnect(SQLite(), 'chinook.db')
  result <- dbGetQuery(conn, query)
  dbDisconnect(conn)
  return(result)
}

run_query('SELECT * FROM customer')
```
Create a show_tables() function that calls the run_query() function to return a
list of all tables and views in the database.

```{r}
show_tables <- function(){
  run_query("SELECT name FROM sqlite_master WHERE type='table'")
}

show_tables()
```
Write a query that returns each genre, with the number of tracks sold in the USA:
in absolute numbers
in percentages

Create a plot to show this data.
```{r}
query <- "
WITH usa_invoice_line AS
  (
  SELECT il.* FROM invoice_line il
  LEFT JOIN invoice i ON i.invoice_id = il.invoice_id
  LEFT JOIN customer c ON c.customer_id = i.customer_id
  WHERE c.country = 'USA'
  ),
  total_tracks_sold_usa as
  (
  SELECT COUNT(*) as total_sold FROM usa_invoice_line
  )

SELECT 
  g.name as genre,
  count(il.invoice_id) as tracks_sold,
  cast(count(il.invoice_id) as FLOAT)/(SELECT total_sold FROM total_tracks_sold_usa) as pct_track_sold
FROM genre as g
LEFT JOIN track tr ON g.genre_id = tr.genre_id
LEFT JOIN usa_invoice_line il ON il.track_id = tr.track_id
GROUP BY 1
ORDER BY 3 DESC
limit 5
"

genre_track_sold = run_query(query)

ggplot(data = genre_track_sold) +
  aes(x = reorder(genre, -pct_track_sold), y = pct_track_sold) +
  geom_bar(stat = 'identity') +
  labs(title = 'Percentage of tracks solds per genre',
       x = 'genre',
       y = 'Percentage of tracks sold')
```
 Write a paragraph that interprets the data and makes a recommendation for the 
 three artists whose albums we should purchase for the store, based on sales of
 tracks from their genres.

let's analise the artists that sell more tracks of rock in the USA


```{r}
query <- "
WITH usa_rock_invoice_line AS
  (
  SELECT il.* FROM invoice_line il
  LEFT JOIN invoice i ON i.invoice_id = il.invoice_id
  LEFT JOIN customer c ON c.customer_id = i.customer_id
  LEFT JOIN track tr ON tr.track_id = il.track_id
  LEFT JOIN genre g ON tr.genre_id = g.genre_id
  WHERE c.country = 'USA' AND g.name = 'Rock'
  )

SELECT
  ar.name AS artist_name,
  sum(i.total) as value_sold
FROM usa_rock_invoice_line AS il
LEFT JOIN invoice i ON i.invoice_id = il.invoice_Id
LEFT JOIN track tr ON il.track_id = tr.track_id
LEFT JOIN album al ON al.album_id = tr.album_id
LEFT JOIN artist ar ON ar.artist_id = al.artist_id
GROUP BY ar.artist_id
ORDER BY 2 DESC;
"

run_query(query)
```

The artist that have the higher values sold are Van Halen, Nirvana and 
The Rolling Stones



Write a query that finds the total dollar amount of sales assigned to each sales 
support agent within the company. Add any extra attributes for that employee 
that you find are relevant to the analysis.

Create a plot of the results of your query.

Write a short statement describing your results, and provide a possible 
interpretation.


```{r}
query <- "
SELECT
  e.first_name || ' ' || e.last_name as employee_name,
  e.hire_date,
  e.country,
  SUM(i.total) total_sold
FROM employee AS e
LEFT JOIN customer AS c ON e.employee_id = c.support_rep_id
LEFT JOIN invoice as i ON c.customer_id = i.customer_id
WHERE e.title = 'Sales Support Agent'
GROUP BY e.employee_id
"

sales_by_employee <- run_query(query)
sales_by_employee
```
 The employee with more sales is the older and the least sales is the newest.
 
 
```{r}
ggplot(data = sales_by_employee) +
  aes(x = reorder(employee_name, -total_sold), y = total_sold) +
  geom_bar(stat = 'identity') +
  labs(title = 'Total Sold by Employee',
       x = 'Employee',
       y = 'Total Sold')
```
Write a query that collates data on purchases from different countries. For each
country, include:
total number of customers;
total value of sales;
average value of sales per customer;
average order value.

Where a country has only one customer, collect them into an "Other" group.
The results should be sorted by the total sales from highest to lowest, with the
"Other" group at the very bottom.

```{r}
query <- "
WITH country_group_resume AS
  (
  SELECT 
    CASE
      WHEN COUNT(DISTINCT c.customer_id) = 1 THEN 'OTHER'
      ELSE c.country
    END AS country_group,
    COUNT(DISTINCT c.customer_id) as total_customers,
    SUM(i.total) as total_sold,
    COUNT(i.invoice_id) as number_sales
  FROM customer c
  LEFT JOIN invoice i ON c.customer_id = i.customer_id
  GROUP BY c.country
  )
  
SELECT
  country_group,
  SUM(total_customers) as total_customers,
  SUM(total_sold) as total_sold,
  SUM(total_sold)/(sum(total_customers)) as avg_value_per_customer,
  SUM(total_sold)/(SUM(number_sales)) as avg_value_per_sale
FROM country_group_resume
GROUP BY 1 
ORDER BY 2 DESC;
"

sales_by_country <- run_query(query)
sales_by_country
```
For each dimension, create a visualization which demonstrates the data we 
collated in the previous step.
You should decide whether the "Other" group is relevant to your analysis and 
make decisions on where to include it (if anywhere) in your visualizations.
Write a few sentences interpreting your data and visualizations, and make one 
or more recommendations to the marketing team on which countries have potential 
for growth.

```{r}
ggplot(data = sales_by_country) +
  aes(x = reorder(country_group, -total_customers), y = total_customers) +
  geom_bar(stat = 'identity') +
  labs(title = 'Customers per Country',
       x = 'Country',
       y = 'Customers') +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = sales_by_country) +
  aes(x = reorder(country_group, -total_sold), y = total_sold) +
  geom_bar(stat = 'identity') +
  labs(title = 'Total Sold per Country',
       x = 'Country',
       y = 'Total Sold') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = sales_by_country) +
  aes(x = reorder(country_group, -avg_value_per_customer), y = avg_value_per_customer) +
  geom_bar(stat = 'identity') +
  labs(title = 'Average Value Spent per Customer per Country',
       x = 'Country',
       y = 'Average Value Spent per Customer') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = sales_by_country) +
  aes(x = reorder(country_group, -avg_value_per_sale), y = avg_value_per_sale) +
  geom_bar(stat = 'identity') +
  labs(title = 'Average Value Spent per Sale per Country',
       x = 'Country',
       y = 'Average Value Spent per Sale') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
There is many countries with only one customer and countries that have English
as official language have the majority of the clients.
The total sold per country follow the same tendency as the number of costumers.
The average spent per customer per country do not have a big variance, except
the Czech Republic that is a outlier, but there is only 2 costumers in this 
country.
The average value spent per sale it almost the same also.



Write a query that categorizes each invoice as either an album purchase or not,
and calculates the following summary statistics:
Number of invoices;
Percentage of invoices.
Write one to two sentences explaining your findings, and making a prospective
recommendation on whether the Chinook store should continue to buy full albums
from record companies.

```{r}
query <- "
WITH resume_invoice as
    (
    SELECT 
        count(t.track_id) tracks_bought, 
        AVG(t.album_id) average_album_id,
        CAST(MIN(t.album_id) as float) min_album_id
    FROM invoice i
    LEFT JOIN invoice_line iv ON iv.invoice_id = i.invoice_id
    LEFT JOIN track t on t.track_id = iv.track_id
    group by i.invoice_id
    ),
tracks_album as
    (
    SELECT 
        album_id, 
        COUNT(track_id) as n_tracks FROM track
    GROUP BY album_id
    )

SELECT 
    SUM(re.tracks_bought) as tracks_sold,
    COUNT(re.tracks_bought) as count_invoice,
    CAST(COUNT(re.tracks_bought) as float)/(SELECT count(*) FROM resume_invoice) as percentage_invoice,
    CAST(SUM(re.tracks_bought) as float)/(SELECT SUM(re.tracks_bought) FROM resume_invoice) as percentage_sold,
    CASE
        WHEN re.average_album_id = al.album_id
        AND
        tracks_bought = n_tracks THEN 'yes'
    ELSE
        'no'
    END is_album
FROM resume_invoice re
LEFT JOIN tracks_album al on al.album_id = re.min_album_id
GROUP BY is_album;
"

sales_by_country <- run_query(query)
sales_by_country
```

The majority of customers buy tracks and no albuns

