---
title: 'Guided Project: Exploring NYC Schools Data'
author: "Artur Jacques Nürnberg"
date: "8/10/2020"
output: html_document
---


```{r}
library(readr)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
```
```{r}
combined <- read_csv('data/combined.csv')
masterfile <- read_tsv('data/masterfile11_gened_final.txt')
masterfile_d75 <- read_tsv('data/masterfile11_d75_final.txt')
```
```{r}
masterfile_select <- masterfile %>% filter(schooltype == 'High School') %>% select(dbn:aca_tot_11)
masterfile_d75_select <- masterfile_d75 %>% filter(schooltype == 'High School') %>% select(dbn:aca_tot_11)
```

Combine the general education and District 75 survey dataframes using the dplyr 
function bind_rows().

```{r}
masterfile_join <- rbind(masterfile_select, masterfile_d75_select) %>%
  rename(DBN = dbn)
masterfile_join <- left_join(combined, masterfile_join, by = 'DBN')
masterfile_join <- masterfile_join %>% select(-school_name, -highschool)
colnames(masterfile_join) <- gsub(' ', '_', colnames(masterfile_join))
```
Create a correlation matrix to look for interesting relationships between pairs 
of variables in the new, combined dataframe you created. Be sure to consider 
missing values and how you will handle them.

```{r}
masterfile_numeric <- masterfile_join %>% select_if(is.numeric)
cor_mat <- cor(masterfile_numeric, use = "pairwise.complete.obs")
cor_sat <- cor_mat %>% as_tibble(rownames = 'variable') %>% select(variable, avg_sat_score)

strong <- filter(cor_sat, avg_sat_score >0.25 | avg_sat_score < 0.25)
strong <- strong[order(-strong$avg_sat_score),]

print(strong)
```

Create scatter plots to explore potentially interesting relationships between 
variables in greater detail. Recall that, if you have more than a few scatter
plots to make, you can write a function to make multiple plots at once.

```{r}
create_graph<- function(x_var, y_var){
  ggplot(data = masterfile_join, aes_string(x = x_var, y = y_var)) +
    geom_point()
}

x <- strong$variable[2:10]
y <- 'avg_sat_score'

map2(x, y, create_graph)
```
Use the tidyr function pivot_longer() to reshape your dataframe so that the 
survey question (like saf_p_11 and com_s_11) and the score (like 8.7 or 9.0) 
are in two separate columns.

```{r}
masterfile_join_gather <- masterfile_join %>% 
  pivot_longer(cols = saf_p_11:com_s_11, names_to = 'survey_question',
               values_to = 'score')
```

response_type: With the values to indicate "parent," "teacher," "student," 
or "total"
metric: With the values to indicate "engagement," "academic expectations," 
"safety and respect," or "communication"

```{r}
masterfile_join_gather <- masterfile_join_gather %>%
  mutate(
    reponse_type = case_when(
      str_sub(survey_question,4, 6) == '_p_' ~ 'parent' ,
      str_sub(survey_question,4, 6) == '_t_' ~ 'teacher' ,
      str_sub(survey_question,4, 6) == '_s_' ~ 'student' ,
      str_sub(survey_question,4, 6) == '_to' ~ 'total' 
    ),
    metric = case_when(
      str_sub(survey_question, 1, 3) == 'saf' ~ 'safety and respect',
      str_sub(survey_question, 1, 3) == 'com' ~ 'communication',
      str_sub(survey_question, 1, 3) == 'aca' ~ 'academic expectations',
      str_sub(survey_question, 1, 3) == 'eng' ~ 'engagement',
    )
  )
```





