---
title: "Efficient Data Analysis Workflow"
author: "Artur Jacques Nürnberg"
date: "7/31/2020"
output: html_document
---

```{r}
library(readr)
library(dplyr)

book_reviews <- read_csv('book_reviews.csv')
```

dim(book_reviews)

colnames(book_reviews)
"
There is 4 columns (book, review, state and price). Book is the name, review the
score, state that was sold and the price.
"
type_col <- c()

```{r}
for (col in colnames(book_reviews)){
   type_col <- c(type_col, typeof(book_reviews[[col]]))
}
print(type_col)

unique_values <- list(1,1,1,1)

count = 1
for (col in colnames(book_reviews)){
  unique_values[[count]] <- unique(book_reviews[[col]])
  count <- count + 1
}

print(unique_values)
```
"
There is five books, five qualities, six states and five prices.
"

```{r}
filtered_books <- filter(book_reviews, is.na(book_reviews$book))

dim(filtered_books)

filtered_books <- filter(book_reviews, is.na(book_reviews$state))

dim(filtered_books)

filtered_books <- filter(book_reviews, is.na(book_reviews$price))

dim(filtered_books)

filtered_books <- filter(book_reviews, is.na(book_reviews$review))
```

"The only column missing values is the review"

filled_books <- filter(book_reviews, !is.na(book_reviews$review))

print(filled_books)

"Was removed 206 rows"
```{r}
filled_books <- filled_books %>% 
  mutate(
    state = case_when(
      state == 'Texas' ~ 'TX',
      state == 'New York' ~ 'NY',
      state == 'California' ~ 'CA',
      state == 'Florida' ~ 'FL',
      TRUE ~ state
    )
  )

unique(filled_books$state)
```

```{r}
filled_books <- filled_books %>%
  mutate(
    review_num = case_when(
      review == 'Excellent' ~ 5,
      review == 'Great' ~ 4,
      review == 'Good' ~ 3,
      review == 'Fair' ~ 2,
      review == 'Poor' ~ 1,
    ),
    is_high_review = if_else(review_num >= 4, TRUE, FALSE)
  )

print(filled_books)
```

# To verify which one is the most profitable book lets check the sum of the price

```{r}
filled_books %>% group_by(book) %>% summarise(profit = sum(price)) %>% arrange(-profit)
```



